Out-of-Sample Comparisons of Overfit Models

### 2014-02-19 paper.tex
Change bibliography database
More revisions to notation and environment section

\subsection{Notation and Forecasting Environment}
We assume the following forecasting environment. There are two
competing linear models that give forecasts for the target,
$y_{t+\h}$:
\[
y_{t+\h} = x_{1t}'\theta_1 + \e_{1,t+h}, \quad\text{and}\quad
y_{t+\h} = x_{2t}'\theta_2 + \e_{2,t+h};
\]
$t = 1,\dots,T-h$, $\h$ is the forecast horizon and the
variables $y_t$, $x_{1t}$, and $x_{2t}$ are all known in period $t$.
The coefficients $\theta_1$ and $\theta_2$ minimize the population
Mean Square Error, so
\begin{equation*}
  \theta_i = \argmin_\theta \sum_{t=1}^{T-\h} \E (y_{t+\h} - x_{it}'\theta)^2,
\end{equation*}
making $\e_{i,t+\h}$ uncorrelated with $x_{i,t}$; $\e_{i,t+\h}$ may exhibit
serial correlation so both of the linear models may be misspecified.
Let $\mathcal{F}_t = \sigma(y_1, x_1, \dots, y_t, x_t)$ be the information
set available to the forecaster in period $t$,
with $x_t$ the vector of all stochastic elements of $x_{1t}$ and $x_{2t}$ after
removing duplicates, and let $\E_t$ and $\var_t$ denote the
conditional mean and variance given $\mathcal{F}_t$.  The first model
uses $K_1$ regressors, and the second uses $K_2$.  Without loss of
generality, assume that $K_1 \leq K_2$. At least one of the models is
overfit, which we represent asymptotically by letting $K_2$ grow with
$T$ quickly enough that $\lim K_2/T$ is positive; $K_1$ may grow with
$T$ as well. Since the number of regressors changes with $T$, a
stochastic array underlies all of our asymptotic theory, but we
suppress that notation to simplify the presentation.

In the settings we are interested in, a forecaster observes the data
$(y_t,x_t)$ for periods 1 through $T$ and divides the observations
into an estimation sample of the first $R$ observations and a test
sample of the remaining $P$ observations. The forecaster then compares
the models' performance over the test sample, which entails
constructing two sequences of forecasts, (using a fixed-window
estimation strategy)
\begin{equation}
\hat y_{i,t+\h} = x_{it}'\hat{\theta}_{it}, \qquad \text{for } i=1,2;
\ t = R+1,\dots,T-h,
\end{equation}
where
\begin{equation}
  \bh{it} = \Big(\sum_{s=1}^{R-\h} x_{is}x_{is}'\Big)^{-1} \sum_{s=1}^{R-\h}
  x_{is} y_{s+\h}, \qquad \text{for } i=1,2;\ t = R+1,\dots, T - \h.\footnote{It
    may not be clear why we are using a $t$ index in $\bh{it}$,
    since $\bh{it} = \bh{iR}$ almost surely for all $t \leq T-h$.
    We need this index because $\bh{it}$ will be defined for $t > T-\h$
    soon, and will not equal $\bh{iR}$ for those values of $t$.}
\end{equation}

The models are then compared by their forecast errors over the test
sample.  There are many statistics that could be used to make this
comparison, but we focus on perhaps the most natural, the \dmw\ \oost\
test \citep{DiM:95,Wes:96}.\footnote{The core insights of our paper
  apply to other \oos\ statistics as well.} This statistic is based on
the difference in the models' loss over the test sample, $\oosA$,
defined as
\begin{equation}
  \oosA \equiv P^{-1} \oosSum{t}{1} D_t
\end{equation}
where
\begin{equation}\label{eq:16}
  D_t = L(y_{t+\h} - x_{1t}'\bh{1t}) - L(y_{t+\h} - x_{2t}'\bh{2t}),
\end{equation}
and $L$ is a known loss function. Now the \oost\ test is defined as
$\sqrt{P} \oosA / \sh$, where $\sh^2$ is an estimator of the
asymptotic variance of $\oosA$. (Possibly a Heteroskedasticity- and
Autocorrelation-Consistent, or \hac, estimator.)

There are several hypotheses that the \oost\ test has been used
for. Most theoretical work, such as \citet{DiM:95}, \citet{Wes:96},
and \cite{Mcc:07}, has focused on using these statistics to test
hypotheses about the pseudo-true values $\theta_1$ and $\theta_2$.  In
particular, that research focuses on testing the null hypothesis that
\begin{equation}
  \E L(y_{t+\h} - x_{1t}'\theta_{1}) = \E L(y_{t+\h} - x_{2t}'\theta_{2}).
\end{equation}
Another possibility is that the forecaster will use one of the models
to make a number of predictions in the future---call the number of
predictions $Q$---in which case the quantity of interest is
\begin{equation}
  \E_T \oosB \equiv Q^{-1} \sum_{t=T+1}^{T+Q} \E_T D_t,
\end{equation}
where $D_t$ is defined in Equation~\eqref{eq:16} but here uses the
full-sample estimates of the models' parameters,
\begin{equation}
  \bh{it} = \Big(\sum_{s=1}^{T-\h} x_{is}x_{is}'\Big)^{-1} \sum_{s=1}^{T-\h}
  x_{is} y_{s+\h}, \qquad \text{for } i=1,2;\ t = T+1,\dots, T + Q.
\end{equation}
If $\E_T \oosB$ is positive, the second model is expected to forecast
better than the first over the next $Q$ periods, and if $\E_T \oosB$
is negative then the first model is better. We use a conditional
expectation because the coefficient estimates in $\oosB$ are
stochastic but known in period $T$, and their values will determine
the performance of the two models.

Under conventional fixed-$K$ asymptotic theory, $\E_T \oosB$ would
converge in probability to the difference in the expected loss
associated with the pseudo-true models,\footnote{Under appropriate
  stationarity, moment, and weak-dependence conditions, obviously.}
\begin{equation}\label{eq:22}
  \E L(y_{t+\h} - x_{1t}'\theta_{1}) - \E L(y_{t+\h} - x_{2t}'\theta_{2}).
\end{equation}
But if $K_2$ increases with $T$ these quantities can have different
limits. For a simple example, assume squared-error loss, let $x_{1,t}
= 1$ for all $t$, and let $(y_{t+\h},x_{2,t})$ be i.i.d. $N(0,
\Sigma)$.  Then
\begin{align*}
  \E_T \oosB - & \big(\E L(y_{t+\h} - x_{1t}'\theta_{1}) - \E L(y_{t+\h} - x_{2t}'\theta_{2})\big) \\
  &= \big(\E_T (y_{T+\h+1} - \bh{1,T})^2 - \E_T (y_{T+\h+1} - x_{T+1}'\bh{2,T})^2\big) -
  \big(\E y_{T+\h+1}^2 - \E (y_{T+\h+1} - x_{T+1}'\theta_2)^2 \big)\\
  &= (\bh{2,t} - \theta_2)' \Sigma_x (\bh{2,t} - \theta_2) + o_p(1).
\end{align*}
This last term has expectation $\var(y_T) \frac{K_2}{T - K_2 - 1}$ and
would converge to zero if $K_2$ were fixed, but does not when $\lim
K_2 / T > 0$. In Section~\ref{sec:oostheory} we show that the \oost\
statistic can estimate $\E_T \oosB$ under our increasing $K$
asymptotics and does not estimate the expected loss in~\eqref{eq:22}.

Our conditional expectation has been studied heavily in
cross-sectional settings with independent observations. In such a
setting, $\E_T \oosB$ is equal to the difference in the models'
\emph{generalization error}, which has been used widely in the machine
learning literature \citep[see][for further discussion]{HTF:08}.
Moreover, with i.i.d. observations, the expectation of $\E_T \oosB$
equals \citepos{Aka:69} Final Prediction Error (\textsc{fpe}). Both
generalization error and \textsc{fpe} are defined by a model's
performance on a new, independent, data set, but, for lack of a better
term, we will call $\E_T \oosB$ the difference in generalization error
for the rest of the paper with hopefully no risk of confusion.

>>>>

\subsection{Notation and Forecasting Environment}
We assume the following forecasting environment. There are two
competing linear models that give forecasts for the target,
$y_{t+\h}$:
\begin{equation}
y_{t+\h} = x_{1t}'\theta_1 + \e_{1,t+h}, \quad\text{and}\quad
y_{t+\h} = x_{2t}'\theta_2 + \e_{2,t+h};
\end{equation}
$t = 1,\dots,T-h$, $\h$ is the forecast horizon and the
variables $y_t$, $x_{1t}$, and $x_{2t}$ are all known in period $t$.
The coefficients $\theta_1$ and $\theta_2$ minimize the population
Mean Square Error, so
\begin{equation}
  \theta_i = \argmin_\theta \sum_{t=1}^{T-\h} \E (y_{t+\h} - x_{it}'\theta)^2,
\end{equation}
making $\e_{i,t+\h}$ uncorrelated with $x_{i,t}$; $\e_{i,t+\h}$ can exhibit
serial correlation so both of the linear models may be misspecified.
Let $\mathcal{F}_t = \sigma(y_1, x_1, \dots, y_t, x_t)$ be the information
set available in period $t$,
with $x_t$ the vector of all stochastic elements of $x_{1t}$ and $x_{2t}$ after
removing duplicates, and let $\E_t$ and $\var_t$ denote the
conditional mean and variance given $\mathcal{F}_t$.  The first model
uses $K_1$ regressors, and the second uses $K_2$.  Without loss of
generality, assume that $K_1 \leq K_2$. At least one of the models is
overfit, which we represent asymptotically by letting $K_2$ grow with
$T$ quickly enough that $\lim K_2/T$ is positive; $K_1$ may grow with
$T$ as well. Since the models change with $T$, a
stochastic array underlies all of our asymptotic theory, but we
suppress that notation to simplify the presentation.

In the settings we are interested in, a forecaster observes the data
$(y_t,x_t)$ for periods 1 through $T$ and divides the observations
into an estimation sample of the first $R$ observations and a test
sample of the remaining $P$ observations. The forecaster then compares
the models' performance over the test sample, which entails
constructing two sequences of forecasts, (using a fixed-window
estimation strategy)
\begin{equation}
\hat y_{i,t+\h} = x_{it}'\hat{\theta}_{it}, \qquad \text{for } i=1,2;
\ t = R+1,\dots,T-h,
\end{equation}
where
\begin{equation}
  \bh{it} = \Big(\sum_{s=1}^{R-\h} x_{is}x_{is}'\Big)^{-1} \sum_{s=1}^{R-\h}
  x_{is} y_{s+\h}, \qquad \text{for } i=1,2;\ t = R+1,\dots, T - \h.\footnote{It
    may not be clear why we are using a $t$ index in $\bh{it}$,
    since $\bh{it} = \bh{iR}$ almost surely for all $t \leq T-h$.
    But $\bh{it}$ will be defined for $t > T-\h$
    soon and will not equal $\bh{iR}$ for those values of $t$.}
\end{equation}

The models are then compared by their forecast errors over the test
sample.  There are many statistics that have been considered in the
literature, but we focus on perhaps the most natural, the \dmw\ \oost\
test \citep{DiM:95,Wes:96}.\footnote{The core insights of our paper
  apply to other \oos\ statistics as well.} This statistic is based on
the difference in the models' loss over the test sample, $\oosA$,
defined as
\begin{equation}
  \oosA \equiv P^{-1} \oosSum{t}{1} D_t
\end{equation}
where
\begin{equation}\label{eq:16}
  D_t = L(y_{t+\h} - x_{1t}'\bh{1t}) - L(y_{t+\h} - x_{2t}'\bh{2t}),
\end{equation}
and $L$ is a known loss function. The \oost\ test is defined as
$\sqrt{P} \oosA / \sh$, where $\sh^2$ is an estimator of the
asymptotic variance of $\oosA$. (Possibly a Heteroskedasticity- and
Autocorrelation-Consistent, or \hac, estimator.)

Most theoretical research, such as \citet{DiM:95}, \citet{Wes:96},
and \cite{Mcc:07}, has focused on using the \oost\ statistic to test
hypotheses about the pseudo-true values $\theta_1$ and $\theta_2$.  In
particular, that research focuses on testing the null hypothesis that
\begin{equation}\label{eq:21}
  \E L(y_{t+\h} - x_{1t}'\theta_{1}) - \E L(y_{t+\h} - x_{2t}'\theta_{2}) = 0.
\end{equation}
But the population quantities in \eqref{eq:21} do not determine which
model is more accurate in finite samples and, under our asymptotics,
in the limit when models are allowed to be overfit. The models'
accuracy will also depend on the specific estimates of $\theta_1$ and
$\theta_2$ used to produce the forecasts. In this setting, where the
forecaster will use one of the models to make a number of predictions
in the future---call the number of predictions $Q$---the quantity of
interest becomes
\begin{equation}
  \E_T \oosB \equiv Q^{-1} \sum_{t=T+1}^{T+Q} \E_T D_t,
\end{equation}
where $D_t$ is defined in Equation~\eqref{eq:16} but here uses the
full-sample estimates of the models' parameters,
\begin{equation}
  \bh{it} = \Big(\sum_{s=1}^{T-\h} x_{is}x_{is}'\Big)^{-1} \sum_{s=1}^{T-\h}
  x_{is} y_{s+\h}, \qquad \text{for } i=1,2;\ t = T+1,\dots, T + Q.
\end{equation}
If $\E_T \oosB$ is positive, the second model is expected to forecast
better than the first over the next $Q$ periods, and if $\E_T \oosB$
is negative then the first model is better. We use a conditional
expectation because the coefficient estimates in $\oosB$ are
stochastic but known in period $T$, and their values will determine
the performance of the two models.

Under conventional fixed-$K$ asymptotic theory, $\E_T \oosB$ would
converge in probability to the difference in the expected loss
associated with the pseudo-true models,\footnote{Under standard
  stationarity, moment, and weak-dependence conditions.}

i.e. the LHS of~\eqref{eq:21}.
\begin{equation}\label{eq:22}
  \E L(y_{t+\h} - x_{1t}'\theta_{1}) - \E L(y_{t+\h} - x_{2t}'\theta_{2}).
\end{equation}
But if $K_2$ increases with $T$ these quantities can have different
limits. For a simple example, assume squared-error loss, let $x_{1,t}
= 1$ for all $t$, and let $(y_{t+\h},x_{2,t})$ be i.i.d. $N(0,
\Sigma)$.  Then
\begin{align*}
  \E_T \oosB - & \big(\E L(y_{t+\h} - x_{1t}'\theta_{1}) - \E L(y_{t+\h} - x_{2t}'\theta_{2})\big) \\
  &= \big(\E_T (y_{T+\h+1} - \bh{1,T})^2 - \E_T (y_{T+\h+1} - x_{T+1}'\bh{2,T})^2\big) -
  \big(\E y_{T+\h+1}^2 - \E (y_{T+\h+1} - x_{T+1}'\theta_2)^2 \big)\\
  &= (\bh{2,t} - \theta_2)' \Sigma_x (\bh{2,t} - \theta_2) + o_p(1).
\end{align*}
This last term has expectation $\var(y_T) \frac{K_2}{T - K_2 - 1}$ and
would converge to zero if $K_2$ were fixed, but does not when $\lim
K_2 / T > 0$. In Section~\ref{sec:oostheory} we show that the \oost\
statistic can estimate $\E_T \oosB$ under our increasing $K$
asymptotics and does not estimate the expected loss associated with
the pseudo-true coefficients.

The conditional expectation $\E_T \oosB$ has been studied heavily in
cross-sectional settings with independent observations. In such a
setting, $\E_T \oosB$ is equal to the difference in the models'
\emph{generalization error}, which has been used widely as a measure
of model accuracy in the machine
learning literature \citep[see][for further discussion]{HTF:08}.
Moreover, with i.i.d. observations, the expectation of $\E_T \oosB$
equals \citepos{Aka:69} Final Prediction Error (\textsc{fpe}). Both
generalization error and \textsc{fpe} are defined by a model's
performance on a new, independent, data set, but, for lack of a better
term, we will call $\E_T \oosB$ the difference in generalization error
for the rest of the paper with hopefully no risk of confusion.
